A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2020-01-27 doc/large_strain.tex
by -> on

In this contribution, we proposed and numerically investigated three different matrix-free implementations of tangent operators for finite-strain elasticity with heterogeneous materials.
To the best of our knowledge, this is the first work that evaluates matrix-free sum factorization techniques on the partial differential equations arising in this case.
Among the three examined algorithms,
the implementation that caches the second-order Kirchhoff stress tensor and evaluates terms in the current configuration ({\color{red}Algorithm \ref{alg:mf_tensor2}}) is faster than
caching only a scalar at quadrature points ({\color{red}Algorithm \ref{alg:mf_scalar}}) or caching the material part of the fourth-order spatial tangent stiffness tensor together with the second-order Kirchhoff stress ({\color{red}Algorithm \ref{alg:mf_tensor4}}).
{\color{red}
Algorithm \ref{alg:mf_tensor2} is recommended when linearization of the adopted material model allows to efficiently implement the action of the material part of the fourth-order spatial tangent stiffness tensor {\color{red}by a second-order symmetric tensor}.
Algorithm \ref{alg:mf_tensor4} does
}
not utilize the specific form of the material part of the stiffness tensor.
{\color{red}
In this case
}
the matrix-free implementation of the tangent operator is somewhat slower but more flexible because it can be applied to any constitutive model which
operates with the material part of the fourth-order spatial tangent stiffness tensor at the quadrature points.
{\color{red}
This implementation is still considerably faster than matrix-based strategies for polynomial degrees higher than one.
}

>>>>

In this contribution, we proposed and numerically investigated three different matrix-free implementations of tangent operators for finite-strain elasticity with heterogeneous materials.
To the best of our knowledge, this is the first work that evaluates matrix-free sum factorization techniques on the partial differential equations arising in this case.
Among the three examined algorithms,
the implementation that caches the second-order Kirchhoff stress tensor and evaluates terms in the current configuration ({\color{red}Algorithm \ref{alg:mf_tensor2}}) is faster than
caching only a scalar at quadrature points ({\color{red}Algorithm \ref{alg:mf_scalar}}) or caching the material part of the fourth-order spatial tangent stiffness tensor together with the second-order Kirchhoff stress ({\color{red}Algorithm \ref{alg:mf_tensor4}}).
{\color{red}
Algorithm \ref{alg:mf_tensor2} is recommended when linearization of the adopted material model allows to efficiently implement the action of the material part of the fourth-order spatial tangent stiffness tensor on a second-order symmetric tensor.
Algorithm \ref{alg:mf_tensor4} does
}
not utilize the specific form of the material part of the stiffness tensor.
{\color{red}
In this case
}
the matrix-free implementation of the tangent operator is somewhat slower but more flexible because it can be applied to any constitutive model which
operates with the material part of the fourth-order spatial tangent stiffness tensor at the quadrature points.
{\color{red}
This implementation is still considerably faster than matrix-based strategies for polynomial degrees higher than one.
}
